### 2021-12-16 input/proposal.tex
address some issues, todo others

\begin{definition}[\label{def:scoped}\underline{S}coped broad\underline{cast} (\NAMEB)]
  When \Process~$x$ scoped \underline{b}roadcasts $b_x(m)$ a
  \underline{m}essage $m$, every correct \process $y$ within a scope
  \underline{r}eceives $r_y(m)$ and \underline{d}elivers it
  $d_y(m)$. The scope depends on the \underline{s}tate $\sigma$ of
  each \process, the \underline{m}etadata $\mu$ piggybacked by each
  message, and a \underline{p}redicate $\phi$ verified from \process
  to \process:
  $(b_x(m) \wedge r_y(m)) \implies \exists \pi_{xy}: \forall z \in
  \pi_{xy}, \phi(\mu_z, \sigma_z)$.
\end{definition}

@@

\begin{definition}[\label{def:forwarding}Forwarding]
  A \process $x$ \underline{f}orwards a message $m$ by
  \underline{s}ending it to its neighbors $y$ \TODO{enriched}
  ($\oplus$) with metadata that depends on its local knowledge
  $\sigma_x$: $f_x(m) \implies \forall \langle x, y\rangle \in E:
  s_{xy}(m \TODO{\oplus} \sigma_x)$.
\end{definition}

@@

Unfortunately, \processes do not share a common global knowledge of
the network state. For \processes to eventually ($\eventually$) reach
consistent partitioning, each source $s$ must send
\texttt{\underline{A}dd} notifications $\alpha_s$ to all \processes
that are closer to it than any other source, along with enough
metadata to allow them to decide of their closest source. A simple
implementation based on Definition~\ref{def:scoped} and
Definition~\ref{def:forwarding} consists in ensuring that each
\process $x$ eventually receives $\alpha_s^{|\Pi_{sx}|}$ where
$\Pi_{sx}$ is the \emph{best} path from any source to it; the best
value $|\Pi_{sx}|$ being built over forwarding along this path.

\begin{theorem}[\label{theo:fb}\underline{F}orwarding of \underline{B}est (FB)
    $\implies\eventually \textup{CP}$]
    % 
    Assuming that each \process $x$ stores its outgoing weights
    ($\forall \langle x, y \rangle \in E: \pi_{xy} \in \sigma_x$), a
    total order on messages based on weights ($m^d_s < m^{d'}_{s'}$
    when $d < d' \vee (d = d' \wedge s< s')$), reliable communication
    links ($s_{xy}(m) \iff \eventually r_{yx}(m)$), \processes eventually reach
    consistent partitioning if each \process delivers its best (\ie
    smallest) message among \underline{r}eceived messages $R_x$, and
    forwards its best message among \underline{d}elivered messages
    $D_x$ such that $f_x(\alpha^d_s) \implies \forall \langle x, y
    \rangle \in E: s_{xy}(\alpha_s^d \oplus ^{w_{xy}})$, \ie by
    accumulating the respective weight of used edges.
\end{theorem}

\begin{proof}
  Whenever a \process~$s$ becomes a source, it broadcasts hence
  delivers its own message $d_s(\alpha_{s}^{|\pi_{ss}|})$. Whatever
  its set of received messages, it acknowledges that it belongs to its
  own partition $P_s$ since $\alpha_{s}^{|\pi_{ss}|} = \min D_s$ and
  it remains forever since $|\pi_{ss}|$ is its greatest lower bound.
  \noindent Such a source forwards its notification to its neighbors.
  Every neighbor eventually receives its notification since
  communication links are reliable. Whatever the order of received
  messages $R_x$ at neighboring \process~$x$, total order ensures that
  it delivers notifications $\alpha_s^{d}$ when $d= |\pi_{ss}| +
  w_{sx} = \min |\pi_{s'x}|$ where $\min |\pi_{s'x}|$ is the lightest
  weight of $x$ to any source $s'$.
  \noindent Among these neighbors, at least those that fulfill the
  latest condition forward their respective notification.  By
  transitivity, the message originating from $s$ reaches all
  \processes that belong to $P_s$ at least through their respective
  lightest path: $\small\smash{\forall y \in V, s, s' \in S: \min
    |\pi_{s y}| < \min |\pi_{s' y}| \implies \eventually
    d_{y}(\alpha_{s}^{\min |\pi_{sy}|})}$. When the system becomes
  quiescent, \ie no \process becomes source anymore, every \process
  eventually acknowledges the partition it belongs to, \ie \processes
  eventually reach consistent partitioning together. In addition, the
  protocol terminates: a \process never delivers hence forwards a
  message after it received, delivered, and forwarded the message of
  its closest source from its lightest path.
\end{proof}

@@

Algorithm~\ref{algo:add} shows the instructions that implement a
\emph{reactive} forwarding of best to reach consistent partitioning in
a static network where \processes never join, crash, nor leave the
system. As soon as a \process receives a better message, it delivers
and forwards it. The trade-off consists in decreasing termination time
at cost of increased traffic.  Figure~\ref{fig:add} illustrates the
behavior of this algorithm on a system comprising 4 \processes $a$,
$b$, $c$, and $d$. Both $a$ and $d$ become sources.  They \NAMEB
notifications $\alpha_a^0$ and $\alpha_d^0$. They initialize their own
state with the lowest value $0$ (see Line~\ref{line:lowestbound}), and
send a message to each of their neighbors by accumulating the
corresponding edge weight (see Line~\ref{line:accumulator}). In
Figure~\ref{fig:addC}, $b$ receives $\alpha_{d}^{1}$. Since it
improves its own partition distance, it keeps it and forwards it to
its neighbors. In Figure~\ref{fig:addD}, $b$ discards
$\alpha_{a}^{2}$, for it does not improve its partition
distance. Neither $c$ nor $d$ will ever acknowledge that Source~$a$
exists.  The protocol lacks of obvious traffic optimization, \eg grey
messages are useless in this scenario. Nevertheless, the system
discards last transiting messages after it reached consistent
partitioning.

@@

Delete operations bring a new notion of order between events, and most
importantly between message deliveries. A \process $s$ that performs a
\texttt{Del} operation implicitly states that all its preceding
\texttt{Add} operations become obsolete. As consequence, their results
-- scoped broadcast in the form of \texttt{\underline{A}dd}
notifications $\alpha_s$ -- should be canceled by the corresponding
\texttt{\underline{D}el} notifications $\delta_s$. We focus on the
last delivery of each \process, since the best is also the last, as
highlighted by Algorithm~\ref{algo:add}.

@@

A naive attempt at implementing DCP resembles echoes in
acoustics. Following the principles of scoped broadcast as stated in
Definition~\ref{def:scoped}, a \process that receives a staleness
notification forwards it only if the latter targets its current
partition. The staleness notifications propagate through the network
as long as they remain in the deleted partition. When they reach the
bordering \processes of the deleted partition, it creates an echo of
bordering partitions that will go backward to fill the gap left open
by removals using forwarding of best (FB) as stated by
Theorem~\ref{theo:fb}.

@@

As a cascading effect in Figure~\ref{fig:problemD}, not only $c$
wrongfully believes that it belongs to $a$'s partition when $a$
already deleted it, but it could contribute to its inconsistency by
blocking farther but up-to-date messages from actual sources.

A first step to avoid staying in inconsistent state is to extend the
behavior of each \process so \processes such as $c$, that would remain
in a wrong partition, use their history of receipt and delivery to
detect the blocking of staleness notifications that can hinder
reaching consistent partitioning. For instance, $c$ detects the
blocking of $\delta_a$ because it acknowledges that its parent $b$
delivered a stale notification $\alpha_c^1$, \ie between the delivery
of $\alpha_c^1$ and the delivery of $\delta_c$, $b$ blocked any other
$\delta$ and therefore $\delta_a$.

>>>>

\begin{definition}[\label{def:scoped}Scoped broadcast]
  When \Process~$x$ scoped \underline{b}roadcasts $b_x(m)$ a
  \underline{m}essage $m$, every correct \process $y$ within a scope
  \underline{r}eceives $r_y(m)$ and \underline{d}elivers it
  $d_y(m)$. The scope depends on the \underline{s}tate $\sigma$ of
  each \process, the \underline{m}etadata $\mu$ piggybacked by each
  message, and a \underline{p}redicate $\phi$ verified from \process
  to \process:
  $(b_x(m) \wedge r_y(m)) \implies \exists \pi_{xy}: \forall z \in
  \pi_{xy}, \phi(\mu_z, \sigma_z)$.
\end{definition}

@@

\begin{definition}[\label{def:forwarding}Forwarding]
  When Node $x$ \underline{f}orwards $f_x(m)$ a message $m$, it
  \underline{s}ends $s_{xy}(m)$ the latter to all its neighbors $y$
  \TODO{enriched} ($\oplus$) with metadata that depends on its local
  knowledge $\sigma_x$:
  $f_x(m) \implies \forall \langle x, y\rangle \in E: s_{xy}(m
  \TODO{\oplus} \sigma_x)$.
\end{definition}

@@

Unfortunately, \processes do not share a common global knowledge of
the network state. For \processes to eventually ($\eventually$) reach
consistent partitioning (CP), each source $s$ must send
\texttt{\underline{A}dd} notifications $\alpha_s$ to all \processes
that are closer to it than any other source, along with enough
metadata to allow them to decide of their closest source. Based on
Definition~\ref{def:scoped} and Definition~\ref{def:forwarding}, this
consists in ensuring that each \process $x$ eventually receives
$\alpha_s^{|\Pi_{sx}|}$ where $\Pi_{sx}$ is the \emph{best} path from
any source to it; the best value $|\Pi_{sx}|$ being built over
forwarding along this path.

\begin{theorem}[\label{theo:fb}\underline{F}orwarding of \underline{B}est (FB)
    $\implies\eventually \textup{CP}$]
    % 
    Assuming that each \process $x$ stores its outgoing weights
    ($\forall \langle x, y \rangle \in E: w_{xy} \in \sigma_x$), a
    total order on messages based on weights ($m^d_s < m^{d'}_{s'}$
    when $d < d' \vee (d = d' \wedge s< s')$), reliable communication
    links ($s_{xy}(m) \iff \eventually r_{yx}(m)$), \processes eventually reach
    consistent partitioning if each \process delivers its best (\ie
    smallest) message among \underline{r}eceived messages $R_x$, and
    forwards its best message among \underline{d}elivered messages
    $D_x$ such that $f_x(\alpha^d_s) \implies \forall \langle x, y
    \rangle \in E: s_{xy}(\alpha_s^d \oplus ^{w_{xy}})$, \ie by
    accumulating the respective weight of used edges.
\end{theorem}

\begin{proof}
  Whenever a \process~$s$ becomes a source, it broadcasts hence
  delivers its own message $d_s(\alpha_{s}^{|\pi_{ss}|})$. Whatever
  its set of received messages, it acknowledges that it belongs to its
  own partition $P_s$ since $\alpha_{s}^{|\pi_{ss}|} = \min D_s$ and
  it remains forever since $|\pi_{ss}|$ is its greatest lower bound.
  \noindent Such a source forwards its notification to its neighbors.
  Every neighbor eventually receives its notification since
  communication links are reliable. Whatever the order of received
  messages $R_x$ at neighboring \process~$x$, total order ensures that
  it delivers notifications $\alpha_s^{d}$ when
  $d= |\pi_{ss}| + w_{sx} = \min |\pi_{s'x}|$ where $\min |\pi_{s'x}|$
  is the lightest weight of $x$ to any source $s'$, $s'$ being $s$ in
  this case.
  \noindent Among these neighbors, at least those that fulfill the
  latest condition forward their respective notification.  By
  transitivity, the message originating from $s$ reaches all
  \processes that belong to $P_s$ at least through their respective
  lightest path: $\small\smash{\forall y \in V, s, s' \in S: \min
    |\pi_{s y}| < \min |\pi_{s' y}| \implies \eventually
    d_{y}(\alpha_{s}^{\min |\pi_{sy}|})}$. When the system becomes
  quiescent, \ie no \process becomes source anymore, every \process
  eventually acknowledges the partition it belongs to, \ie \processes
  eventually reach consistent partitioning together. In addition, the
  protocol terminates: a \process never delivers hence forwards a
  message after it received, delivered, and forwarded the message of
  its closest source from its lightest path.
\end{proof}

@@

Algorithm~\ref{algo:add} shows the instructions that implement a
\emph{reactive} forwarding of best to reach consistent partitioning in
a static network where \processes never join, crash, nor leave the
system. As soon as a \process receives a better message, it delivers
and forwards it. The trade-off consists in decreasing termination time
at cost of increased traffic.  Figure~\ref{fig:add} illustrates the
behavior of this algorithm on a system comprising 4 \processes $a$,
$b$, $c$, and $d$. Both $a$ and $d$ become sources.  They scoped broadcast
notifications $\alpha_a^0$ and $\alpha_d^0$. They initialize their own
state with the lowest value $0$ (see Line~\ref{line:lowestbound}), and
send a message to each of their neighbors by accumulating the
corresponding edge weight (see Line~\ref{line:accumulator}). In
Figure~\ref{fig:addC}, $b$ receives $\alpha_{d}^{1}$. Since it
improves its own partition distance, it keeps it and forwards it to
its neighbors. In Figure~\ref{fig:addD}, $b$ discards
$\alpha_{a}^{2}$, for it does not improve its partition
distance. Neither $c$ nor $d$ will ever acknowledge that Source~$a$
exists.  The protocol lacks of obvious traffic optimization, \eg grey
messages are useless in this scenario. Nevertheless, the system
discards last transiting messages after it reached consistent
partitioning.

@@

Delete operations bring a new notion of order between events, and most
importantly between message deliveries. A \process $s$ that performs a
\texttt{Del} operation implicitly states that all its preceding
\texttt{Add} operations become obsolete. As consequence, their results
-- scoped broadcast in the form of \texttt{\underline{A}dd}
notifications $\alpha_s$ -- should be canceled by the corresponding
\texttt{\underline{D}el} notifications of staleness $\delta_s$. We
focus on the last delivery of each \process, since the best is also
the last, as highlighted by Algorithm~\ref{algo:add}. \TODO{More about
  useless $\alpha$ ?}

@@

A naive attempt at implementing DCP resembles echoes in acoustics: a
sound propagates in the air before crashing into surrounding walls to
finally come back altered. Following the principles of scoped
broadcast as stated in Definition~\ref{def:scoped}, a \process $x$
that receives a staleness notification $\delta_s$ forwards it only if
the latter targets its current partition $\alpha_s$. The staleness
notifications propagate through the network as long as they remain in
the deleted partition. When they reach the bordering \processes of the
deleted partition, it creates an echo of bordering partitions that
will go backward to fill the gap left open by removals using
forwarding of best (FB) as stated by Theorem~\ref{theo:fb}.

@@

As a cascading effect in Figure~\ref{fig:problemD}, not only $c$
wrongfully believes that it belongs to $a$'s partition when $a$
already deleted it, but it could contribute to its inconsistency by
not forwarding farther but up-to-date messages from actual sources.

A first step to avoid staying in inconsistent state is to extend the
behavior of each \process so \processes such as $c$, that would remain
in a wrong partition, use their history of receipt and delivery to
detect the blocking of staleness notifications that can hinder
reaching consistent partitioning. For instance, $c$ detects the
blocking of $\delta_a$ because it acknowledges that its parent $b$
delivered a stale notification $\alpha_c^1$, \ie between the delivery
of $\alpha_c^1$ and the delivery of $\delta_c$, $b$ blocked any other
$\delta$ and therefore $\delta_a$.  \TODO{Must take action because its
  parent has forwarded something already stale so previous cannot be
  good.}
